---
title: "Homework_week9_functions"
author: "Roshan"
date: "2023-04-13"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```


# The 3 functions.

### Function 1: **roundpenguins** 
### Function 2: **penguinviolin** 
### Function 3: **redtext**




# Load libraries

```{r}
library(tidyverse)
library(dplyr)
library(here)
library(palmerpenguins)
library(kableExtra)
library(ggplot2)
```


# Creating function 1

```{r}
## creating roundpenguins as a new function containing 3 sub-functions
roundpenguins <- function(penguins) {  
  penguins %>%              ## calling 'penguins' dataset as argument for mutate function
    mutate(bill_depth_mm = round(bill_depth_mm),  ## rounds bill_depth column to integers
           bill_length_mm = round(bill_length_mm)) %>%  ## round bill_length col to integers
    na.omit() ## removes any rows containing empty fields in any columns
}

## using 'roundpenguins' function to clean penguin data columns
roundedpenguins <- roundpenguins(penguins)  
view(roundedpenguins)
```

# Example of function benefits

The table below shows the printed 'head' of the roundedpenguin data after function
roundpenguins was executed carrying out 3 independent functions. Rounding column
"bill depth" and "bill length" to an integer using the mutate function and removing NA's
using na.omit
```{r}
head(roundedpenguins)
```

# Creating function 2

```{r}


penguinviolin <- function(penguins) {   ## creating new function with penguin argument for data to produce a penguin plot
  ggplot(data = penguins,aes(x = bill_length_mm, 
          y = body_mass_g,
          color = species,
          fill = species))+    
          geom_violin()+## using violin plot 
  labs(x = "Bill Length (mm)",     # Axis labels 
       y = "Body Mass (g)")  }



p1 <- penguinviolin(penguins) ## creating output "p1" using the new "penguinviolin" function
p1 ## displaying penguin violin plot
  


```

# Creating function 3

```{r}

redtext <- function() {
  theme(plot.title = element_text(color = "red"),
        axis.text = element_text(color = "red"),
        axis.title = element_text(color = "red"))
  }

```

# Showing example of function 3

```{r}

### so just copying code for a ggplot and on the last line the redtext function is added
ggplot(data = penguins,aes(x = bill_length_mm, 
          y = body_mass_g,
          color = species,
          fill = species))+    
          geom_violin()+## using violin plot 
  labs(x = "Bill Length (mm)",     # Axis labels 
       y = "Body Mass (g)") +
  redtext()  ## here the redtext function now changes x and y axis to red and axis labels also.

p1+redtext()    ## these functions work together nicely :)


```






